---
title: "Penguin Data Visualization"
author: "Tom"
date: "2023-10-04"
output: html_document
---

# Introduction

In this document, we will explore the Palmer Penguins dataset using various data visualization techniques in R. We will use the `ggplot2` package to create plots and examine the relationship between flipper length and body mass among penguins, while considering different aspects of the data such as species and sex.

For more details on using R Markdown, [click here](http://rmarkdown.rstudio.com).

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

# Loading the Data

```{r loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(palmerpenguins)
```
# Load the penguins dataset
```{r}
data(penguins)
```
# Display a glimpse of the dataset
```{r}
glimpse(penguins)
```

# Scatter Plots
#### Scatter Plot: Flipper Length vs. Body Mass

We can create a basic scatter plot to visualize the relationship between flipper length and body mass among penguins:

``` {r}
# Scatter plot of flipper length vs. body mass
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(color = "purple")
```

#### Scatter Plot with Species Differentiation

Now, let's add color differentiation for different penguin species:

``` {r}
# Scatter plot with color-coded points for species
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
``` 

#### Scatter Plot with Species and Shape Differentiation

To further enhance the plot, we can use both color and shape to differentiate the points by species:

``` {r}
# Scatter plot with color-coded and shape-coded points for species
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species))
```

#### Faceted Scatter Plots by Sex

Now, let's create faceted scatter plots by sex, considering the flipper length and body mass:

``` {r}
# Faceted scatter plots by sex with color-coded and shape-coded points for species
penguins %>%
  drop_na(sex) %>%
  ggplot(aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(aes(color=species,
                 shape=species)) +
  facet_wrap(~sex)
```

# Conclusion

In this document, we explored the Palmer Penguins dataset using various scatter plots to visualize the relationship between flipper length and body mass among penguins. We considered differentiation by species and sex to gain insights into the data.